---
title: "Homework 2 - data wrangling with the tidyverse"
author: "Runxiang Liu"
date: "14/2/2023"
output:
  md_document:
    toc: yes
    variant: markdown_github
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

# Homework 2 - data wrangling with the `tidyverse`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Which NBA player scored the most points in 1991?

```{r echo=FALSE}
library(tidyverse)

# Read in the data
data <- read.csv("Seasons_Stats_NBA.csv")
data %>% 
  filter(Year == 1991) %>% 
  arrange(desc(PTS)) %>% 
  select(Player) %>% 
  head(1)
```

## Which player had the best free throw percentage from the year 2000 to the most recent year in the data?

```{r echo=FALSE}
data %>% 
  filter(Year >= 2000) %>% 
  arrange(desc(FT.)) %>% 
  select(Player) %>% 
  head(1)
```

## Rename the variable `pos` to `position`.
```{r}
data %>% 
  rename(Position = Pos) %>% head(5)
```

## Use this variable to create two variables that are called `first_position` and `second_position`. Hint: separate by splitting the position variable in two.

```{r}
data$first_position <- str_split(data$position, "-") %>% unlist %>% .[[1]]
data$second_position <- str_split(data$position, "-") %>% unlist %>% .[[2]]
```
## Create two new datasets.
### a new dataset from the original dataset that includes all data except the age variable (be sure to give this dataset a new name).

```{r}
data %>% 
  select(-Age) -> data1
head(data1, 5)
```

### a new dataset from the original dataset that includes the year, the player name, and age.
```{r}
data %>% 
  select(Year, Player, Age) -> data2
head(data2, 5)
```

## add a new column to both datasets called mergeid that includes a sequence of numbers beginning with a 1 in the first row of the data and ending with the total number of rows in the last row of the data
```{r}
data1$mergeid <- seq(1,nrow(data1),1)

data2$mergeid <- seq(1,nrow(data2),1)
```

## Join the two datasets from question (6) together to recreate the original dataset plus the new merge id.
```{r}
new_data <- merge(data1, data2, by = "mergeid")
head(new_data, 5)
```

## Subset the original dataset to 1995. Group the data by year and team name and then summarize the average number of points per team. Arrange from most to least points.
```{r}
data %>% filter(Year==1995) %>% group_by(Year, Tm) %>% summarize(avg_pts = mean(PTS)) %>% arrange(desc(avg_pts)) -> data3
head(data3, 5)
```

## Reshape the data in the previous question into a wide format using the tidyr package. Create a wide dataset that keeps year in a single column, but spreads team names to multiple individual columns with each column delineating points per team in 1995.

```{r}
data3 %>% spread(Tm, avg_pts) -> data4
head(data4, 5)
```

## Now return the data to a long (tidy) format by moving teams back into a single column and points in a single column.

```{r}
data4 %>% gather(Tm, avg_pts, -Year) -> data5
head(data5, 5)
```


